---
title: "Assignment 2"
author: "Spencer Moon"
date: "10/28/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Data was loading using the following code:

```{r results='hide', message=FALSE, warning=FALSE}
library(tidyverse)

redwine <- read.table('redwine.txt', header = TRUE, sep = "\t", na.strings = 'NA')
```


# Problem 1


The averages of RS and SD without the missing values are as follows:

```{r}
rs_avg <- mean(redwine$RS, na.rm = TRUE)
sd_avg <- mean(redwine$SD, na.rm = TRUE)

# Print averages
paste('RS average:', round(rs_avg, digits = 2))
paste('SD average:', round(sd_avg, digits = 2))
```


# Problem 2


The coefficients of a linear model between SD and FS are as follows:

```{r}
# Creat SD and FS vectors
SD.obs <- redwine$SD[is.na(redwine$SD) == FALSE]
FS.obs <- redwine$FS[is.na(redwine$SD) == FALSE]  

# Build simple regression and print coefficients
SDFS_fit <- lm(SD.obs ~ FS.obs)
coefficients(SDFS_fit)
```


# Problem 3


Missing values of SD were imputed using the following code:

```{r}
# Create vector of estimated SD
FS_fill <- data.frame(FS.obs = redwine$FS[is.na(redwine$SD) == TRUE])
SD_est <- data.frame(predict(SDFS_fit, FS_fill))

# Create function for imputing missing SD values
estimp <- function(field, est)
{
  missing	<- is.na(field)
  n.missing <- sum(missing)
  field.obs <- field[!missing]
  imputed	<- field
  for (i in 1:n.missing)
  {
    imputed[missing][i] = est[i,]
  }
  return(imputed)
}

# Use function and print new RS average
redwine$SD <- estimp(redwine$SD, SD_est)
paste('SD new average:', round(mean(redwine$SD), digits = 2))
```


# Problem 4


Missing values of RS were imputed using the following code:

```{r}
# Create function for imputing missing RS values
avgimp <- function(field, avg)
{
  missing	<- is.na(field)
  n.missing <- sum(missing)
  field.obs <- field[!missing]
  imputed	<- field
  imputed[missing] <- avg
  return(imputed)
}

# Use function and print new RS average
redwine$RS <- avgimp(redwine$RS, rs_avg)
paste('RS new average:', round(mean(redwine$RS), digits = 2))
```


# Problem 5

Below is the multiple linear regression model:

```{r}
winemodel <- lm(QA ~ ., redwine)
coefficients(winemodel)
```


# Problem 6


Below is the summary of the model:

```{r}
summary(winemodel)
```

Based on the summary shown above, the PH variable is least likely to be related to QA as it has the highest *p*-value compared to other variables.


# Problem 7

```{r}
library(DAAG)

validation <- CVlm(data = redwine, m = 3, form.lm = winemodel, printit = FALSE)
```

Error is 0.43.


# Problem 8


The average and standard deviation are shown below:

```{r}
ph_avg <- mean(redwine$PH)
ph_std <- sd(redwine$PH)

paste('PH average:', round(ph_avg, digits = 2))
paste('PH standard deviation:', round(ph_std, digits = 2))
```


Below are the dimensions of *redwine* and *redwine2* as well as the number of rows removed:

```{r}
redwine2 <- subset(redwine , PH < ph_avg + 3*ph_std & PH > ph_avg - 3*ph_std)
dim(redwine)
dim(redwine2)
paste("Number of rows removed:", dim(redwine)[1] - dim(redwine2)[1])
```


# Problem 9 


Below is the new model:

```{r}
winemodel2 <- lm(QA ~ ., redwine2)
summary(winemodel2)
```

The new model is worse. Even though we have a very small increase in R squared, we have less significant coefficients. Based on having the smallest *p*-values, variables VA, CH, SD, SU, and AL are most likely to be related to QA.