---
title: "Lab Exercise 1 - Problem 3"
author: "Spencer Moon"
date: "10/15/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Before building the regression model, the data was loaded with the following code:

```{r results='hide', message=FALSE, warning=FALSE}
library(tidyverse)

labdata <- read_tsv("labdata.txt")
```


### Part A


Below is the linear model for the lab data:

```{r}
reg <- lm(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8, labdata)
summary(reg)
```


### Part B


```{r}
attach(labdata)
plot(x1, y)
#plot(x2, y)
#plot(x3, y)
#plot(x4, y)
#plot(x5, y)
#plot(x6, y)
#plot(x7, y)
#plot(x8, y)
```

Variable x1 can be used in a piecewise regression model as the scatter plot of x1 against y clearly shows a kink.


### Part C

```{r}
avg = mean(x1)
avg

reg.piece <- lm(y ~ (x1<avg)*x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8, labdata)
summary(reg.piece)
```

Above summary shows that the piecewise regression model of the lab data is indeed better because the R-squared value is signficantly greater than the original regression model.