---
title: "Assignment 3"
author: "Spencer Moon"
date: "11/13/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 1

The data was loaded using the following code: 

```{r results='hide', message=FALSE, warning=FALSE}
# Load necessary libraries for this exercise
library(expm)

# Import data
markov <- read.table('markov100.txt')
markov <- as.matrix(markov)
```


### Part A

```{r}
# Create state vector with first element as 1 (currently at State 1)
a <- c(rep(0, dim(markov)[1]))
a[1] <- 1

# Find probability of being in State 5 after 10 transitions
trans <- 10
result <- a %*% (markov %^% trans)
result[5]
```

The probability of being in State 5 after 10 transitions from State 1 is 4.5%.


### Part B

```{r}
# Create state vector with first element as 1 (currently at State 1)
a <- c(rep(0, dim(markov)[1]))
prob <- 1/3
a[c(1,2,3)] <- prob

# Find probability of being in State 5 after 10 transitions
trans <- 10
result <- a %*% (markov %^% trans)
result[10]
```

The probability of being in State 10 after 10 transitions from State 1, 2, or 3 with equal probability is 8.3%.


### Part C

```{r}
# Replace the last row of (P transpose - I) by a vector of ones
Q <- t(markov) - diag(dim(markov)[1])
Q[dim(markov)[1],] <- c(rep(1, dim(markov)[1]))

# Solve for pi
rhs <- c(rep(0, dim(markov)[1]))
rhs[dim(markov)[1]] <- 1
Pi <- solve(Q) %*% rhs
Pi[1]
```

The steady state probability of State 1 is 1.3%.


### Part D

```{r}
B <- markov[-100, -100]
Q <- diag(dim(markov)[1] - 1) - B
e = c(rep(1, dim(markov)[1] - 1))
m = solve(Q) %*% e
m[1]
```

The mean first passage time from State 1 to State 100 is 254.94.


# Problem 2

The data was loaded using the following code:

```{r}
web <- read.table('webtraffic.txt', header = TRUE)
```


### Part A

```{r}
# Create vector of column sums
Traffic <- as.vector(colSums(web))

# Create 9 by 9 matrix from vector
Traffic <- matrix(Traffic, nrow = 9, ncol = 9, byrow = TRUE)
Traffic
```


### Part B

```{r}
# Change one matrix element
Traffic[9,1] = 1000
P <- Traffic/rowSums(Traffic)
P
```


### Part C

```{r}
# Replace the last row of (P transpose - I) by a vector of ones
Q <- t(P) - diag(dim(P)[1])
Q[dim(P)[1],] <- c(rep(1, dim(P)[1]))

# Solve for pi
rhs <- c(rep(0, dim(P)[1]))
rhs[dim(P)[1]] <- 1
Pi <- solve(Q) %*% rhs
Pi
```


### Part D

```{r warning = FALSE}
# Create average time vector
time <- c(0.1, 2, 3, 5, 5, 3, 3, 2, 0)

# Calculate average time
sum(Pi * time)
```

The average time sepnt on the website is 2.3 minutes.


### Part E

```{r}
# Adjust for additional links
Traffic[2,6] <- .3 * Traffic[2,3]
Traffic[2,3] <- .7 * Traffic[2,3]
Traffic[2,7] <- .2 * Traffic[2,4]
Traffic[2,4] <- .8 * Traffic[2,4]

# Recalculate probability matrix
P <- Traffic/rowSums(Traffic)

# Replace the last row of (P transpose - I) by a vector of ones
Q <- t(P) - diag(dim(P)[1])
Q[dim(P)[1],] <- c(rep(1, dim(P)[1]))

# Solve for pi
rhs <- c(rep(0, dim(P)[1]))
rhs[dim(P)[1]] <- 1
Pi2 <- solve(Q) %*% rhs

# Check variance differences of Pi and Pi2
var(Pi)
var(Pi2)
```

The variance for Pi2 is slightly lower, which indicates that the traffic has become more balanced with the new links.
